---
title: "ST558-Project 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = FALSE)
library(tidyverse)
```

# Introduction

TBD: Mark

# Data

```{r}
# Import News
news <- read_csv("OnlineNewsPopularity/OnlineNewsPopularity.csv")

# Split by channel
news <- news %>%
  pivot_longer(starts_with("data_channel_is_"), 
               names_prefix = "data_channel_is_",
               names_to = "channel") %>%
  filter(value == 1) %>%
  select(-value) %>%
  split(.$channel) 
```

# Summarizations

Each group member is responsible for producing some summary statistics (means, sds, contingency tables, etc.) and for producing at least three graphs (each) of the data.

# Modeling

The first group member should fit a random forest model and the second group member should
fit a boosted tree model. Both models should be chosen using cross-validation.

Prior to the models fit using linear regression, the first group member should provide a short
but thorough explanation of the idea of a linear regression model.

Prior to each ensemble model, you should provide a short but reasonably thorough explanation
of the ensemble model you are using (so one for each group member).

## Random Forest

TBD: Jovanni

## Boosted Tree

TBD: Mark

# Comparison

This can be done by one group member and the automation done by the other (see below).

# Automation

This should be done by the group member that doesn’t automate the comparison of models
part.
