---
title: "ST558-Project 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = FALSE)
library(tidyverse)
```

# Introduction

TBD: Mark

# Data

```{r}
# Import News
news <- read_csv("OnlineNewsPopularity/OnlineNewsPopularity.csv")

# Split by channel
news <- news %>%
  pivot_longer(starts_with("data_channel_is_"), 
               names_prefix = "data_channel_is_",
               names_to = "channel") %>%
  filter(value == 1) %>%
  select(-value) %>%
  split(.$channel) 
```
Let's create the file names and params for the render code
```{r echo = FALSE, eval=FALSE}
## Create file names
data_channels <- names(news)
output_file <- paste0(data_channels, ".html")

## Create list for each data channel with just the channel name param.
params <- lapply(data_channels, FUN = function(x){list(data_channel = x)})

## Put into df
reports <- tibble(output_file, params)
```
```{r echo=FALSE, eval=FALSE}
## Render code
library(rmarkdown)
apply(reports, MARGIN = 1,
      FUN = function(X){
        render(input = "ST558-Project2.Rmd",
               output_file = x[[1]],
               params = x[[2]])
      })
```
Let's return the dataset based on the specified data channel param for each report
```{r}
data_channel <- params$data_channel
news <- news$data_channel
```
# Summarizations
```{r}
##
##correlations <- t(cor(news$shares, news[, 2:60]))
ggplot(data = news, aes(shares, LDA_02)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000) 

ggplot(data = news, aes(shares, avg_negative_polarity)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000) 

ggplot(data = news, aes(shares, average_token_length)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000) 

ggplot(data = news, aes(shares, max_negative_polarity)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000) 

ggplot(data = news, aes(shares, kw_max_avg)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000)

ggplot(data = news, aes(shares, LDA_03)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000)


ggplot(data = news, aes(shares, kw_avg_avg)) +
  geom_point(aes(color = as.factor(data_channel_is_world))) +
  xlim(0, 15000) 
```
Get days of week in a single column and plot visuals based on weekday for shares
```{r}
newsDaysSub <- news %>% select(shares, weekday_is_monday, weekday_is_tuesday, weekday_is_wednesday, weekday_is_thursday, weekday_is_friday, weekday_is_saturday, weekday_is_sunday)

newsDaysSubLong <- pivot_longer(data = newsDaysSub, cols = c(weekday_is_monday, weekday_is_tuesday, weekday_is_wednesday, weekday_is_thursday, weekday_is_friday, weekday_is_saturday, weekday_is_sunday), names_to = "Day")

ggplot(newsDaysLong, aes(shares)) +
  geom_boxplot(aes(fill=Day)) +
  scale_fill_discrete(labels = c("Friday", "Monday",  "Saturday","Sunday", "Thursday","Tuesday", "Wednesday")) +
  xlim(0, 15000)
```
Each group member is responsible for producing some summary statistics (means, sds, contingency tables, etc.) and for producing at least three graphs (each) of the data.

# Modeling

The first group member should fit a random forest model and the second group member should
fit a boosted tree model. Both models should be chosen using cross-validation.

Prior to the models fit using linear regression, the first group member should provide a short
but thorough explanation of the idea of a linear regression model.

Prior to each ensemble model, you should provide a short but reasonably thorough explanation
of the ensemble model you are using (so one for each group member).

## Random Forest

TBD: Jovanni

## Boosted Tree

TBD: Mark

# Comparison

This can be done by one group member and the automation done by the other (see below).

# Automation

This should be done by the group member that doesn’t automate the comparison of models
part.
